---
title: "Class 1 Survey Analysis"
author: "Fariha Chowdhury"
date: "2024-09-14"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### **INSTRUCTIONS**

#### **Complete all tasks 1-7 using reproducible coding practices (annotation of all code). Use R syntax to solve all problems (i.e. do not manually change values/variables in datasets or check them manually to make sure things are correct except as a double or triple check). Please cut and paste each problem before each R markdown chunk and annotate your code within the chunks. Compare and consolidate your answers. Select one member of your group to walk the class through how each problem was solved. Elect one member of your group to turn in your consolidated code with answers.**

## 

## 

##### 1. Import class one survey data from our Github site (https://github.com/kijohnson/ADA-2024) (The dataset is called 'Class 1 Survey Fall 2024_di.csv'), calling the R dataframe that you create *C1survey*.

```{r}
# Import the C1survey dataset from Github
pacman::p_load(dplyr, readr)
C1survey<-read_csv("https://raw.githubusercontent.com/kijohnson/ADA-2024/main/Class%201%20(Getting%20started%20and%20RR%20I)/Lab/Class%201%20Survey%20Fall%202024_di.csv")
```

##### 2. Determine: **a.** the number of observations (i.e. the number of people who filled out the survey) and **b.** the number of variables in the dataframe.    

```{r}
# a and b find out the number of observations and variables in C1survey dataset
dim(C1survey)
# there are 29 observations and 27 variables
str(C1survey)
```

##### 3. **a.** Rename the column variables to something shorter and that is descriptive of what the variable is about (for example *like_dogs* for the 'Do you like dogs?' question variable) and **b.** Display that they are renamed.

```{r}
# using names function to rename the 27 variables in order
names(C1survey)[1:27]<-c("id", "like_cats", "like_dogs", "have_desert", "slogan", "fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby", "program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length", "top_three","public_health_interest", "fav_num", "bday", "bmonth", "country", "state", "city", "highest_educ_level")

# b. show the renamed variables
colnames(C1survey)
```

##### 4. Write code to determine and display the number of factor, integer, numerical, and character variables there are in the *C1survey* dataset.

```{r}
#using table and sapply functions 23 character and 4 integer
table(sapply(C1survey, class)) 
summary(C1survey)
```

##### 5. **a.** Using code, check the *bday* and *bmonth* variables for any unusual or missing values. If you find any, **b.** describe what you will do with the unusual values in a comment before or immediately following the code that does it, and **c.** after you have cleaned up any unusual values, find the median bday and month. *Note: please make sure that your bday and month variables are numeric before doing statistics

```{r}
# a. looking at values for bday and bmonth
C1survey$bday
C1survey$bmonth

# b. recoding bday with non numeric values

# bday recoding method 
class(C1survey$bday) # check that it is coded as character class
C1survey$bday<-ifelse(C1survey$bday == "May 21-report 21", "21", C1survey$bday)
C1survey$bday<-ifelse(C1survey$bday == "March 31st", "31", C1survey$bday)
C1survey$bday # check values 

# convert to numeric for statistics
C1survey$bday<-as.numeric(C1survey$bday)
table(C1survey$bday, useNA = "always") # check values 

# bmonth recoding
class(C1survey$bmonth)

C1survey$bmonth[C1survey$bmonth=="May 21-report 5"]<-"5"
C1survey$bmonth[C1survey$bmonth=="February"]<-"2"
C1survey$bmonth[C1survey$bmonth=="September"]<-"9"
C1survey$bmonth[C1survey$bmonth=="July"]<-"7"
C1survey$bmonth[C1survey$bmonth=="March"]<-"3"
C1survey$bmonth # check values 

# convert to numeric for statistics
C1survey$bmonth<-as.numeric(C1survey$bmonth)
C1survey$bmonth # check values 

# c. calculate medians for bday and bmonth 
median(C1survey$bday, na.rm=TRUE) # 14 (biased because of 0, which is not a value) How could we better design a survey to ensure this doesn't happen?
median(C1survey$bmonth, na.rm=TRUE) # 7
```

##### 6. **a.** Create a new variable called *bseason* that gives the season according to Northern Meteorological season in which respondents were born (winter = 12/1 to 2/29, spring = 3/1 to 5/31, summer = 6/1 to 8/31, fall = 9/1 to 11/30). **b**. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct **c**. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code. 

```{r}
# a.
# using dplyr
C1survey <- C1survey %>%
  mutate(bseason=case_when(bmonth %in% c(12,1,2) ~"Winter",
                           bmonth %in% c(3,4,5) ~"Spring",
                           bmonth %in% c(6,7,8) ~"Summer",
                           bmonth %in% c(9,10,11) ~"Fall"))
table(C1survey$bmonth, C1survey$bseason) # check variable

C1survey$bseason_f<-factor(C1survey$bseason, levels=c("Winter", "Spring", "Summer", "Fall"))


# b. 
table(C1survey$bmonth, C1survey$bseason)
table(C1survey$bmonth, C1survey$bseason_f)

# c.
tab<-addmargins(table(C1survey$bmonth, C1survey$bseason, useNA = "always"), 1)
tab
tab<-addmargins(table(C1survey$bmonth, C1survey$bseason_f, useNA = "always"), 1)
tab


# Births by season: Fall = 4 Winter = 9 Spring = 8 Summer = 8
```

##### 7. Pick your favorite variable or variables to analyze, come up with a question you want to answer with that variable or variables, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. #'s).

```{r}
#Let’s say I choose the variable like_dogs to analyze, which indicates whether respondents like dogs. The question I want to answer is: How many people in the survey like dogs?
# I'll analyze the variable *like_dogs* to see how many people like dogs.
# First, I check the levels of the variable to understand the possible responses.
table(C1survey$like_dogs)

# Now, I'll count how many people responded 'Yes' to liking dogs.
# Assuming 'Yes' is a response in the *like_dogs* variable.
like_dogs_yes <- sum(C1survey$like_dogs == "Yes")

# I'll display the result.
like_dogs_yes

```
